Cluster Dynamics with HETDEX at z < 0.5 - I: Simulated Performance, Mass Recovery, and Limits to Cosmology

### 2017-01-27 paper1.tex
even more edits

We investigate the accuracy of the halo mass distribution by comparing the cumulative number density of halos above a mass ($M_{200c}$) threshold to the halo mass function (HMF) of \cite{Tinker2008}. We calculate the HMF at central redshifts of 0.1, 0.2, and 0.4 using {\sc HMFcalc} \citep{Murray2013} and compare it to galaxy clusters in a redshift window of $\Delta z\pm0.01$. We find a very good agreement between the predicted HMF and the observed distribution of clusters.

@@

\subsection{Mock Observations}\label{sec: observations}
When selecting galaxies from the Buzzard catalog we assume an observation for all galaxies laying within a turquoise, IFU square in Figure~\ref{fig: ifu layout}. In practice, this is achieved by three dither positions at each pointing. Because of the 1:3 fill factor, many galaxies will not receive spectroscopic coverage, including galaxies that lie between IFUs within a pointing, as well as galaxies between pointings. We use a pointing pattern which differs slightly from the expected final HETDEX observing pattern. See Figure~\ref{fig: ifu layout} for an illustration of the difference. To cover the 398.49 \degsq\ field of the Buzzard catalog we require 5370 pointings where 0.015 \degsq\ of each pointing is covered by an IFU. The total area of the sky covered by an IFU is 80.80 \degsq\ which gives a filling factor of 1/4.65, slightly (about 3\%) decreased from the expected filling factor of 1/4.5.

@@

We follow the maximum likelihood method of \cite{Walker2006} to estimate the line-of-sight velocity dispersion (LOSVD). Using the previously calculated LOSVs ($v_i$ in the following discussion), we maximize the probability function:
\begin{equation}\label{eq: jointGaussian}
P(\{v_1, ..., v_N\})=\displaystyle\prod_{i=1}^{N}\frac{1}{\sqrt{2\pi(\sigma_i^2+\sigma_{1D}^2)}}\exp\biggl[-\frac{1}{2}\frac{(v_i-\langle u \rangle)^2}{(\sigma_i^2+\sigma_{1D}^2)}\biggr]
\end{equation}
	where $\sigma_{1D}$, $\langle u \rangle$, and $\sigma_i$ is the LOSVD, the average radial velocity and the error (which we have assumed to be normally distributed) on the individual $v_i$ measurements respectively. To maximize Equation~\ref{eq: jointGaussian}, we employ a Monte Carlo Markov Chain (MCMC) sampler ({\sc emcee}\footnote{\url{http://dan.iel.fm/emcee/current/}}; \citealt{Foreman-Mackey2013}) which is based on affine-invariant ensemble sampler (see \citealt{Goodman2010} for details on affine-invariant samplers). We draw twenty thousand samples from the posterior probability distribution using simple priors, $\langle\mu\rangle$ lies between the maximum and minimum LOSV and $0< \sigma_{1D} < 1400$ \kms, for each cluster. We set the upper limit on the LOSVD as the LOSVD corresponding to a $10^{16}$ \Msol\ cluster at $z\sim0.0$, higher mass than any expected cluster in the Buzzard catalogs. When the complete sampling of the posterior probability distribution is not used, the final LOSVD is quoted as the median value of the posterior probability distribution with 68\% error bars defined as the square root of second moment of the same distribution, the standard deviation.

@@

Figure \ref{fig: recovery} shows the recovery fraction of member galaxies, the number of observed galaxies divided by the number of actual galaxies ($N_{obs}/N_{True}$), as a function of both redshift and cluster mass. As expected, the Targeted observing strategy where the individual clusters are targeted through several dithers to ensure near complete coverage, performs significantly better than the Survey observing strategy across all redshifts and cluster masses. By definition, with Perfect knowledge, recovery fraction would be unity across all redshifts and cluster masses where clusters exist.

@@

For the recovery fraction as a function of true cluster mass (x-direction), we find a general decrease in the number of member galaxies with increasing cluster mass. This  is entirely a result of the flux limited survey (see previous paragraph). Because there are few high mass ($M_{200c}>5\times10^{14}$ \Msol) clusters, many of which are at moderately high redshift, the higher redshift cluster members suffer from the limiting apparent magnitude and suppress the recovery fraction at fixed mass. If we were to limit the Survey to $z<0.2$ we find the recovery fraction of clusters, across all masses, increases substantially, and we find a much more consistent detection fraction across all masses. However, doing so limits severely the number of clusters at masses greater than $10^{14}$ \Msol because of the limited survey volume.

@@

For the Targeted and Survey observations the power law predicted cluster masses give $\mu_{bias} =0.135\pm{0.003}$ dex, $\sigma_{bias} = 0.370\pm{0.002}$ dex and $\mu_{bias} =0.148\pm{0.008}$ dex, $\sigma_{bias} = 0.324\pm{0.006}$ dex, respectively. So for the clusters that we detect with Survey observations, we obtain similar levels of accuracy as to the Targeted observations, on the average. This does not mean that the Survey observations cannot be improved by Targeted observations. In fact, when comparing only the galaxies which overlap between the two samples the bias and scatter of the Targeted observations is significantly decreased as more cluster member galaxies are detected, better sampling the LOSVD probability distribution function. The Targeted observations perform similarly on the average because many lower mass clusters are included in the sample, increasing the bias of the overall sample.

>>>>

We investigate the accuracy of the halo mass distribution by comparing the cumulative number density of halos above a mass ($M_{200c}$) threshold to the halo mass function (HMF) of \cite{Tinker2008}. We calculate the HMF at central redshifts of 0.1, 0.2, and 0.4 using {\sc HMFcalc} \citep{Murray2013} and compare it to galaxy clusters in a redshift window of $\Delta z\pm0.01$. In each of the three test cases we use a $\chi^2$-test, using clusters with masses $10^{10} - 10^{15}$ \msol. We find very good agreement between the predicted HMF and the observed distribution of clusters with the minimum $\chi^2$ statistic $\sim0.004$ and a corresponding p-value $>0.95$.

@@

\subsection{Mock Observations}\label{sec: observations}
When selecting galaxies from the Buzzard catalog we assume an observation for all galaxies laying within a turquoise, IFU square in Figure~\ref{fig: ifu layout}. In practice the individual IFUs have a fill factor of 1:3, so a three position dither pattern at each pointing insures complete coverage. Galaxies which do not receive spectroscopic coverage include galaxies that lie between the IFUs within a single pointing, as well as galaxies between the pointings. We use a pointing pattern which is designed to mimic that of HETDEX but differs slightly from the expected final HETDEX observing pattern. See Figure~\ref{fig: ifu layout} for an illustration of the difference.To illustrate the difference between the pointing patterns, in order to cover the 398.49 \degsq\ field of the Buzzard catalog we require 5370 pointings where 0.015 \degsq\ of each pointing is covered by an IFU (\ie, a turquoise square in Figure~\ref{fig: ifu layout}). After completing the survey, the total area of the sky covered by an IFU is 80.80 \degsq\ which gives a filling factor of 1/4.65, slightly (about 3\%) decreased from the expected filling factor of 1/4.5.

@@

We follow the maximum likelihood method of \cite{Walker2006} to estimate the line-of-sight velocity dispersion (LOSVD). Using the previously calculated LOSVs ($v_i$ in the following discussion), we maximize the probability function:
\begin{equation}\label{eq: jointGaussian}
P(\{v_1, ..., v_N\})=\displaystyle\prod_{i=1}^{N}\frac{1}{\sqrt{2\pi(\sigma_i^2+\sigma_{1D}^2)}}\exp\biggl[-\frac{1}{2}\frac{(v_i-\langle u \rangle)^2}{(\sigma_i^2+\sigma_{1D}^2)}\biggr]
\end{equation}
where $\sigma_{1D}$, $\langle u \rangle$, and $\sigma_i$ is the LOSVD, the average radial velocity and the error (which we have assumed to be normally distributed) on the individual $v_i$ measurements respectively. To maximize Equation~\ref{eq: jointGaussian}, we employ a Monte Carlo Markov Chain (MCMC) sampler ({\sc emcee}\footnote{\url{http://dan.iel.fm/emcee/current/}}; \citealt{Foreman-Mackey2013}) which is based on affine-invariant ensemble sampler (see \citealt{Goodman2010} for details on affine-invariant samplers). We draw twenty thousand samples from the posterior probability distribution using simple priors, $\langle\mu\rangle$ lies between the maximum and minimum LOSV and $0< \sigma_{1D} < 1400$ \kms, for each cluster. We set the upper limit on the LOSVD as the LOSVD corresponding to a $10^{16}$ \Msol\ cluster at $z\sim0.0$, higher mass than any expected cluster in the Buzzard catalogs. The resulting estimates of LOSVD are not sensitive to these particular choices of priors. When the complete sampling of the posterior probability distribution is not used, the final LOSVD is quoted as the median value of the posterior probability distribution with 68\% error bars defined as the square root of second moment of the same distribution, the standard deviation.

@@

Figure \ref{fig: recovery} shows the recovery fraction of member galaxies, the number of observed galaxies divided by the number of actual galaxies ($N_{obs}/N_{True}$), as a function of both redshift and cluster mass. As expected, the Targeted observing strategy where the individual clusters are targeted through several dithers to ensure near complete coverage, performs significantly better than the Survey observing strategy across all redshifts and cluster masses. The Targeted observations detect $\sim10\%$ of all possible clusters whereas the Survey observations detect only $\sim1\%$. By definition, with Perfect knowledge, recovery fraction would be unity across all redshifts and cluster masses where clusters exist.

@@

For the recovery fraction as a function of true cluster mass (x-direction), we find a general decrease in the number of member galaxies with increasing cluster mass. This  is entirely a result of the flux limited survey (see previous paragraph). Because there are few high mass ($M_{200c}>5\times10^{14}$ \Msol) clusters, many of which are at moderately high redshift, the higher redshift cluster members suffer from the limiting apparent magnitude and suppress the recovery fraction at fixed mass. If we were to limit the Survey to $z<0.2$ we find the recovery fraction of clusters, at fixed masses, increases 40\%-50\%, and we find a much more consistent detection fraction across all masses. However, doing so limits severely the number of clusters at masses greater than $10^{14}$ \Msol because of the limited survey volume.

@@

For the Targeted and Survey observations the power law predicted cluster masses give $\mu_{bias} =0.135\pm{0.003}$ dex, $\sigma_{bias} = 0.370\pm{0.002}$ dex and $\mu_{bias} =0.148\pm{0.008}$ dex, $\sigma_{bias} = 0.324\pm{0.006}$ dex, respectively. So for the clusters that we detect with Survey observations, we obtain similar levels of accuracy as to the Targeted observations, on the average. This does not mean that the Survey observations cannot be improved by Targeted observations. In fact, when comparing only the galaxies which overlap between the two samples the bias and scatter of the Targeted observations is significantly decreased as more cluster member galaxies are detected, owing to better sampling of the LOSVD probability distribution function. The Targeted observations perform similarly on the average because many lower mass clusters are included in the sample, increasing the bias of the overall sample.
